---
title: "DBs Comparative"
output: html_document
---

#Load libraries
```{r}
library(ggplot2)
library(forcats)
library(ggpubr)
```

#Load files
```{r}
#Metadata
metadata=read.csv('./Data/metadata_all_def.csv',header = T, sep='\t')
##Change United States for United\nStates
metadata$country[metadata$country=="United States"]<-"United\nStates"
metadata$host_lifestyle[metadata$host_lifestyle=="Periurban shantytown"]<-"Periurban\nshantytown"
#Load data files for K2 Results
nt_k2_results=read.csv('./Data/nt_K2_results.txt',header = T, sep='\t')
ntE_k2_results=read.csv('./Data/nt_enrich-fMAG_K2_results.txt',header = T, sep='\t')
#Load Bracken count tables
nt_brack_table_SP=data.frame(t(read.csv('./Data/p_nt_SP_count_bracken.txt',sep='\t',row.names = 1)))
nt_brack_table_GEN=data.frame(t(read.csv('./Data/p_nt_GEN_count_bracken.txt',sep='\t',row.names = 1)))
ntE_brack_table_SP=data.frame(t(read.csv('./Data/p_nt_enrich_fMAG_SP_count_bracken.txt',sep='\t',row.names = 1)))
ntE_brack_table_GEN=data.frame(t(read.csv('./Data/p_nt_enrich_fMAG_GEN_count_bracken.txt',sep='\t',row.names = 1)))
```

#K2 Results

##Process tables
```{r}
##Process nt table
#Process tables to get only tax level percentages
nt_k2_results_taxp=nt_k2_results[c("Sample_Name", "Domain","Phylum","Class","Order","Family","Genus","Species")]
nt_k2_results_taxp_m=reshape2::melt(nt_k2_results_taxp,id="Sample_Name")
#Add colum with database name
nt_k2_results_taxp_m$DataBase <- 'NT_DB'
#Change additional name columns
names(nt_k2_results_taxp_m)[names(nt_k2_results_taxp_m)=="variable"] <- "Taxon"
names(nt_k2_results_taxp_m)[names(nt_k2_results_taxp_m)=="value"] <- "Percentage"

##Process ntE table
#Process tables to get only tax level percentages
ntE_k2_results_taxp=ntE_k2_results[c("Sample_Name", "Domain","Phylum","Class","Order","Family","Genus","Species")]
ntE_k2_results_taxp_m=reshape2::melt(ntE_k2_results_taxp,id="Sample_Name")
#Add colum with database name
ntE_k2_results_taxp_m$DataBase <- 'NT-MS-fMAG_DB'
#Change additional name columns
names(ntE_k2_results_taxp_m)[names(ntE_k2_results_taxp_m)=="variable"] <- "Taxon"
names(ntE_k2_results_taxp_m)[names(ntE_k2_results_taxp_m)=="value"] <- "Percentage"

##Combine both tables
k2_tax_results=rbind(nt_k2_results_taxp_m,ntE_k2_results_taxp_m)
```

##Classification capacity by taxons
```{r}
#Create boxplot
p1=ggplot(data=k2_tax_results, aes(x=DataBase, y=Percentage))+geom_boxplot(aes(fill=DataBase),outlier.colour ="red")
#Format graph
p1=p1+xlab("")+ylab("Reads assignment(%)")
p1=p1+facet_wrap(~ factor(Taxon, levels = c("Domain","Phylum","Class","Order","Family","Genus","Species")) , 
                 scales="free",ncol=7,as.table = FALSE)
p1=p1+theme(axis.text.x = element_blank(),axis.ticks.x=element_blank())
p1=p1+coord_cartesian(ylim = c(0, 100))+theme(strip.text.x = element_text(size = 11))
#Add mean to graph
p1=p1+stat_summary(aes(group=DataBase),fun=mean, geom="point", shape=25, size=2, color="black", fill="red")
#Show result
print(p1)
```

```{r}
#Genus Stats
k2_tax_results_GEN=subset(k2_tax_results, Taxon=="Genus")
class_nt=k2_tax_results_GEN[k2_tax_results_GEN$DataBase=='NT_DB',]
mean(class_nt$Percentage,na.rm=TRUE)
sd(class_nt$Percentage,na.rm = TRUE)
summary(class_nt$Percentage)

class_nt_Ef=k2_tax_results_GEN[k2_tax_results_GEN$DataBase=='NT-MS-fMAG_DB',]
mean(class_nt_Ef$Percentage,na.rm=TRUE)
sd(class_nt_Ef$Percentage,na.rm = TRUE)
summary(class_nt_Ef$Percentage)
```

```{r}
#SP Stats
k2_tax_results_SP=subset(k2_tax_results, Taxon=="Species")
class_nt=k2_tax_results_SP[k2_tax_results_SP$DataBase=='NT_DB',]
mean(class_nt$Percentage,na.rm=TRUE)
sd(class_nt$Percentage,na.rm = TRUE)
summary(class_nt$Percentage)

class_nt_Ef=k2_tax_results_SP[k2_tax_results_SP$DataBase=='NT-MS-fMAG_DB',]
mean(class_nt_Ef$Percentage,na.rm=TRUE)
sd(class_nt_Ef$Percentage,na.rm = TRUE)
summary(class_nt_Ef$Percentage)
```

##Classification capacity by metadata GENUS Level
```{r}
#Filter k2 results table
k2_tax_results_GEN=subset(k2_tax_results, Taxon=="Genus")

#Merge counts and metadata
meta_cross=merge(metadata,k2_tax_results_GEN,by.x = "run_accession",by.y = "Sample_Name")

#By Lifestyle
#Create boxplot
p2=ggplot(data=meta_cross, aes(x=DataBase, y=Percentage))+geom_boxplot(aes(fill=DataBase),outlier.colour ="red")
#Format graph
p2=p2+xlab("")+ylab("Reads assignment(%)")
p2=p2+facet_wrap(~ host_lifestyle, scales="free",ncol=3)+theme(axis.text.x = element_blank(),axis.ticks.x=element_blank())
p2=p2+coord_cartesian(ylim = c(0, 100))+theme(strip.text.x = element_text(size = 11))
#Add mean to graph
p2=p2+stat_summary(aes(group=DataBase),fun=mean, geom="point", shape=25, size=2, color="black", fill="red")
#Show result
print(p2)

#By Country
#Create boxplot
p3=ggplot(data=meta_cross, aes(x=DataBase, y=Percentage))+geom_boxplot(aes(fill=DataBase),outlier.colour ="red")
#Format graph
p3=p3+xlab("")+ylab("Reads assignment(%)")
p3=p3+facet_wrap(~ country, scales="free",ncol=6)+theme(axis.text.x = element_blank(),axis.ticks.x=element_blank())
p3=p3+coord_cartesian(ylim = c(10, 100))+theme(strip.text.x = element_text(size = 11))
#Add mean to graph
p3=p3+stat_summary(aes(group=DataBase),fun=mean, geom="point", shape=25, size=2, color="black", fill="red")
#Show result
print(p3)
```

##Classification capacity by metadata SP Level
```{r}
#Filter k2 results table
k2_tax_results_SP=subset(k2_tax_results, Taxon=="Species")

#Merge counts and metadata
meta_cross=merge(metadata,k2_tax_results_SP,by.x = "run_accession",by.y = "Sample_Name")

#By Lifestyle
#Create boxplot
p4=ggplot(data=meta_cross, aes(x=DataBase, y=Percentage))+geom_boxplot(aes(fill=DataBase),outlier.colour ="red")
#Format graph
p4=p4+xlab("")+ylab("Reads assignment(%)")
p4=p4+facet_wrap(~ host_lifestyle, scales="free",ncol=3)+theme(axis.text.x = element_blank(),axis.ticks.x=element_blank())
p4=p4+coord_cartesian(ylim = c(0, 100))+theme(strip.text.x = element_text(size = 11))
#Add mean to graph
p4=p4+stat_summary(aes(group=DataBase),fun=mean, geom="point", shape=25, size=2, color="black", fill="red")
#Show result
print(p4)

#By Country
#Create boxplot
p5=ggplot(data=meta_cross, aes(x=DataBase, y=Percentage))+geom_boxplot(aes(fill=DataBase),outlier.colour ="red")
#Format graph
p5=p5+xlab("")+ylab("Reads assignment(%)")
p5=p5+facet_wrap(~ country, scales="free",ncol=6)+theme(axis.text.x = element_blank(),axis.ticks.x=element_blank())
p5=p5+coord_cartesian(ylim = c(10, 100))+theme(strip.text.x = element_text(size = 11))
#Add mean to graph
p5=p5+stat_summary(aes(group=DataBase),fun=mean, geom="point", shape=25, size=2, color="black", fill="red")
#Show result
print(p5)
```

##Combine all graphs in one K2 results plot/ Supplementary Figure 3
```{r}
#Add titles
p1=p1+ggtitle("Kraken 2 Taxon Level Results")
p2=p2+ggtitle("Kraken 2 Genus Level")
p4=p4+ggtitle("Kraken 2 Species Level")
#Arrange figures
sfigure3 <- ggarrange(p1,ggarrange(p2,p4,labels = c("B","C"),ncol = 2,legend="none",heights = c(0.5),widths=c(0.5)),
                        ggarrange(p3,p5,labels = c("",""),ncol = 2,legend="none",heights = c(0.5),widths=c(0.5)),
                        labels = c("A","",""),nrow =3 ,common.legend = TRUE,legend = "bottom")
ggsave("./Supplementary/Supplementary_Figure3.pdf",plot=sfigure3,units = "mm",height = 120, width = 200,scale = 2,dpi=600)
```

#Bracken Results

##Get total reads per table
```{r}
#Get total reads per sample table(we cut also use ntE_k2_results)
total_count_samples=nt_k2_results[c("Sample_Name","Total")]
```

##Bracken results at GENUS Level

###Process tables
```{r}
#Process count tables to classified data

#nt bracken GEN count table
##Get classified counts per sample
nt_brack_counts_GEN=as.data.frame(apply(nt_brack_table_GEN, 2, function(x) sum(x)))
names(nt_brack_counts_GEN)[1]<-"Class_Counts"
##Merge with total counts per sample
nt_brack_GEN=merge(nt_brack_counts_GEN,total_count_samples,by.x = "row.names",by.y = "Sample_Name")
##Get percentage of classified reads per sample
nt_brack_GEN$Percentage <- as.vector(t(nt_brack_GEN["Class_Counts"]/nt_brack_GEN["Total"]*100))
##Add colum with database name
nt_brack_GEN$DataBase <- 'NT_DB'

#ntE bracken GEN count table
##Get classified counts per sample
ntE_brack_counts_GEN=as.data.frame(apply(ntE_brack_table_GEN, 2, function(x) sum(x)))
names(ntE_brack_counts_GEN)[1]<-"Class_Counts"
##Merge with total counts per sample
ntE_brack_GEN=merge(ntE_brack_counts_GEN,total_count_samples,by.x = "row.names",by.y = "Sample_Name")
##Get percentage of classified reads per sample
ntE_brack_GEN$Percentage <- as.vector(t(ntE_brack_GEN["Class_Counts"]/ntE_brack_GEN["Total"]*100))
##Add colum with database name
ntE_brack_GEN$DataBase <- 'NT-MS-fMAG_DB'

##Combine both tables
brack_GEN_results=rbind(nt_brack_GEN,ntE_brack_GEN)
```

###General Classification capacity for all samples
```{r}
#Create boxplot
p6=ggplot(data=brack_GEN_results, aes(x=DataBase, y=Percentage))+geom_boxplot(aes(fill=DataBase),outlier.colour ="red")
#Format graph
p6=p6+xlab("")+ylab("Reads assignment(%)")
p6=p6+theme(axis.text.x = element_blank(),axis.ticks.x=element_blank())
p6=p6+coord_cartesian(ylim = c(0, 100))
#Add mean to graph
p6=p6+stat_summary(aes(group=DataBase),fun=mean, geom="point", shape=25, size=2, color="black", fill="red")
#Show result
print(p6)
```

```{r}
class_nt=brack_GEN_results[brack_GEN_results$DataBase=='NT_DB',]
mean(class_nt$Percentage,na.rm=TRUE)

class_nt=brack_GEN_results[brack_GEN_results$DataBase=='NT-MS-fMAG_DB',]
mean(class_nt$Percentage,na.rm=TRUE)
```


###Classification capacity by metadata GENUS Level
```{r}
#Merge counts and metadata
meta_cross=merge(metadata,brack_GEN_results,by.x = "run_accession",by.y = "Row.names")

#By Lifestyle
#Create boxplot
p7=ggplot(data=meta_cross, aes(x=DataBase, y=Percentage))+geom_boxplot(aes(fill=DataBase),outlier.colour ="red")
#Format graph
p7=p7+xlab("")+ylab("Reads assignment(%)")
p7=p7+facet_wrap(~ host_lifestyle, scales="free",ncol=3)+theme(axis.text.x = element_blank(),axis.ticks.x=element_blank())
p7=p7+coord_cartesian(ylim = c(0, 100))+theme(strip.text.x = element_text(size = 11))
#Add mean to graph
p7=p7+stat_summary(aes(group=DataBase),fun=mean, geom="point", shape=25, size=2, color="black", fill="red")
#Show result
print(p7)

#By Country
#Create boxplot
p8=ggplot(data=meta_cross, aes(x=DataBase, y=Percentage))+geom_boxplot(aes(fill=DataBase),outlier.colour ="red")
#Format graph
p8=p8+xlab("")+ylab("Reads assignment(%)")
p8=p8+facet_wrap(~ country, scales="free_y")+theme(axis.text.x = element_blank(),axis.ticks.x=element_blank())
p8=p8+coord_cartesian(ylim = c(10, 100))+theme(strip.text.x = element_text(size = 11))
#Add mean to graph
p8=p8+stat_summary(aes(group=DataBase),fun=mean, geom="point", shape=25, size=2, color="black", fill="red")
#Show result
print(p8)
```

##Bracken results at SP Level

###Process tables
```{r}
#Process count tables to classified data

#nt bracken GEN count table
##Get classified counts per sample
nt_brack_counts_SP=as.data.frame(apply(nt_brack_table_SP, 2, function(x) sum(x)))
names(nt_brack_counts_SP)[1]<-"Class_Counts"
##Merge with total counts per sample
nt_brack_SP=merge(nt_brack_counts_SP,total_count_samples,by.x = "row.names",by.y = "Sample_Name")
##Get percentage of classified reads per sample
nt_brack_SP$Percentage <- as.vector(t(nt_brack_SP["Class_Counts"]/nt_brack_SP["Total"]*100))
##Add colum with database name
nt_brack_SP$DataBase <- 'NT_DB'

#ntE bracken GEN count table
##Get classified counts per sample
ntE_brack_counts_SP=as.data.frame(apply(ntE_brack_table_SP, 2, function(x) sum(x)))
names(ntE_brack_counts_SP)[1]<-"Class_Counts"
##Merge with total counts per sample
ntE_brack_SP=merge(ntE_brack_counts_SP,total_count_samples,by.x = "row.names",by.y = "Sample_Name")
##Get percentage of classified reads per sample
ntE_brack_SP$Percentage <- as.vector(t(ntE_brack_SP["Class_Counts"]/ntE_brack_SP["Total"]*100))
##Add colum with database name
ntE_brack_SP$DataBase <- 'NT-MS-fMAG_DB'

##Combine both tables
brack_SP_results=rbind(nt_brack_SP,ntE_brack_SP)
```

###General Classification capacity for all samples
```{r}
#Create boxplot
p9=ggplot(data=brack_SP_results, aes(x=DataBase, y=Percentage))+geom_boxplot(aes(fill=DataBase),outlier.colour ="red")
#Format graph
p9=p9+xlab("")+ylab("Reads assignment(%)")
p9=p9+theme(axis.text.x = element_blank(),axis.ticks.x=element_blank())
p9=p9+coord_cartesian(ylim = c(0, 100))
#Add mean to graph
p9=p9+stat_summary(aes(group=DataBase),fun=mean, geom="point", shape=25, size=2, color="black", fill="red")
#Show result
print(p9)
```

```{r}
class_nt=brack_SP_results[brack_SP_results$DataBase=='NT_DB',]
mean(class_nt$Percentage,na.rm=TRUE)

class_nt=brack_SP_results[brack_SP_results$DataBase=='NT-MS-fMAG_DB',]
mean(class_nt$Percentage,na.rm=TRUE)
```

###Classification capacity by metadata SP Level
```{r}
#Merge counts and metadata
meta_cross=merge(metadata,brack_SP_results,by.x = "run_accession",by.y = "Row.names")

#By Lifestyle
#Create boxplot
pA=ggplot(data=meta_cross, aes(x=DataBase, y=Percentage))+geom_boxplot(aes(fill=DataBase),outlier.colour ="red")
#Format graph
pA=pA+xlab("")+ylab("Reads assignment(%)")
pA=pA+facet_wrap(~ host_lifestyle, scales="free",ncol=3)+theme(axis.text.x = element_blank(),axis.ticks.x=element_blank())
pA=pA+coord_cartesian(ylim = c(0, 100))+theme(strip.text.x = element_text(size = 11))
#Add mean to graph
pA=pA+stat_summary(aes(group=DataBase),fun=mean, geom="point", shape=25, size=2, color="black", fill="red")
#Show result
print(pA)

#By Country
#Create boxplot
pB=ggplot(data=meta_cross, aes(x=DataBase, y=Percentage))+geom_boxplot(aes(fill=DataBase),outlier.colour ="red")
#Format graph
pB=pB+xlab("")+ylab("Reads assignment(%)")
pB=pB+facet_wrap(~ country, scales="free_y")+theme(axis.text.x = element_blank(),axis.ticks.x=element_blank())
pB=pB+coord_cartesian(ylim = c(10, 100))+theme(strip.text.x = element_text(size = 11))
#Add mean to graph
pB=pB+stat_summary(aes(group=DataBase),fun=mean, geom="point", shape=25, size=2, color="black", fill="red")
#Show result
print(pB)
```

##Combine country and life_style metadata boxplots / Figure 1
```{r}
#Add titles for p6 and p9
p6=p6+ggtitle("Bracken Genus Level")
p9=p9+ggtitle("Bracken Species Level")
#Arrange figures
figure1 <- ggarrange(ggarrange(p6,p9,labels = c("A","B"),ncol = 2,legend="none",heights = c(0.5),widths=c(0.5)),
                        ggarrange(p7,pA,labels = c("",""),ncol = 2,legend="none",heights = c(0.5),widths=c(0.5)),
                        ggarrange(p8,pB,labels = c("",""),ncol = 2,legend="bottom",common.legend = TRUE,heights = c(0.5),widths=c(0.5)),
                        labels = c("","",""),nrow =3 ,common.legend = TRUE,legend = "bottom",heights = c(0.5,0.5,1))
ggsave("./Figures/Figure1.pdf",plot=figure1,units = "mm",height = 210, width = 195,scale = 1,dpi=600)
```









